---
title: "Lab 02 - Plastic waste"
author: "Reilly"
date: "4/28/2022"
output: html_document
---

## Load packages and data

```{r setup, message=FALSE}
# Load the tidyverse package
library(tidyverse)
library(ggplot2)
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv(here::here("data/plastic-waste.csv"))
```

## Exercises

### Exercise 1

Continents like Europe and North America have high plastic waste per capita, with one country in North America having almost 4 plastic waste per capita. Africa has the highest number of countries producing the lowest amount of plastic waste per capita compared to other continents. 

```{r plastic-waste-continent}
# insert code here
ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap, fill = continent, color = continent) +
  geom_histogram(binwidth = 0.2) +
  geom_density(aes(y = after_stat(count)), alpha = .5)
```

### Exercise 2

Density curves show the distribution of countries in each continent and where they fall for their plastic waste per capita, and individual points allow certainty in where these countries fall. Box plots are not able to show specific points, only the averages and general distributions. 

```{r plastic-waste-density}
# insert code here
ggplot(na.omit(plastic_waste)) +
  aes(x = continent,
      y = plastic_waste_per_cap,
      fill = continent,
      color = continent) +
  geom_jitter(height = 0,
              width = .4) +
  geom_boxplot(color = "black",
               alpha = .5) +
  ggdist::stat_slab(height = .3,
                    color = "black",
                    size = .2,
                    alpha = .5,
                    position = position_nudge(y = .3))
```

### Exercise 3

Remove this text, and add your answer for Exercise 3 here.

```{r}
ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap,
      y = mismanaged_plastic_waste_per_cap,
      color = continent) +
  geom_point()
```


### Exercise 4

Remove this text, and add your answer for Exercise 4 here.

```{r plastic-waste-violin}
# insert code here
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r plastic-waste-mismanaged}
# insert code here
```

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.

```{r plastic-waste-mismanaged-continent}
# insert code here
```

### Exercise 7

Remove this text, and add your answer for Exercise 7 here.

```{r plastic-waste-population-total}
# insert code here
```

```{r plastic-waste-population-coastal}
# insert code here
```

### Exercise 8

Remove this text, and add your answer for Exercise 8 here.

```{r recreate-viz}
# insert code here
```

## Pro-Tips

### Excercise 3

Try this :D

```
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()+
  geom_boxplot(width=.3, fill="green") +
  stat_summary(fun.y=median, geom="point") 
```

### Exercise 5 

Helpful reference:http://www.sthda.com/english/wiki/ggplot2-themes-and-background-colors-the-3-elements



