---
title: "Lab 02 - Plastic waste"
author: "Reilly"
date: "4/28/2022"
output: html_document
---

## Load packages and data

```{r setup, message=FALSE}
# Load the tidyverse package
library(tidyverse)
library(ggplot2)
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv(here::here("data/plastic-waste.csv"))
```

## Exercises

### Exercise 1

Continents like Europe and North America have high plastic waste per capita, with one country in North America having almost 4 plastic waste per capita. Africa has the highest number of countries producing the lowest amount of plastic waste per capita compared to other continents. 

```{r plastic-waste-continent}
# insert code here
ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap, fill = continent, color = continent) +
  geom_histogram(binwidth = 0.2) +
  geom_density(aes(y = after_stat(count)), alpha = .5)
```

### Exercise 2

Density curves show the distribution of countries in each continent and where they fall for their plastic waste per capita, and individual points allow certainty in where these countries fall. Box plots are not able to show specific points, only the averages and general distributions. 

```{r plastic-waste-density}
# insert code here
ggplot(na.omit(plastic_waste)) +
  aes(x = continent,
      y = plastic_waste_per_cap,
      fill = continent,
      color = continent) +
  geom_jitter(height = 0,
              width = .4) +
  geom_boxplot(color = "black",
               alpha = .5) +
  ggdist::stat_slab(height = .3,
                    color = "black",
                    size = .2,
                    alpha = .5,
                    position = position_nudge(y = .3))
```

### Exercise 3.1-3.2

There appears to be a somewhat positive relationship between plastic waste per capita and mismanaged plastic waste.

Upon coloring the points by continent, continents like Asia, Oceania, and South America seem to have countries that have a high mismanaged plastic waste per capita. Europe appears to have the least mismanaged plastic waste per capita while having a high plastic waste per capita. 

```{r plastic-waste-mismanaged-plastic-waste}
ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap,
      y = mismanaged_plastic_waste_per_cap,
      color = continent) +
  geom_point()
```


### Exercise 3.3.1

There appears to be a moderate association between total population and plastic waste per capita. Although, there are some outliers that tend to skew the trend line, there is a negative association between the two variables.

```{r plastic-waste-total population}
# insert code here
ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap,
      y = total_pop) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE)
```

### Exercise 3.3.2

THe relationship between coastal population and plastic waste per capita is slightly stronger than total population's relationship. There is a negative association between the two variables. 

```{r plastic-waste--coastal}
# insert code here
ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap,
      y = coastal_pop) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE)
```

### Exercise 17.4.4

There appears to be a slight positive correlation between plastic waste and coastal population proportion. There are several African countries that have a low coastal pop proportion to plastic waste relationship, while Oceania have a higher relationship. 

```{r recreate-viz}
# insert code here
new_plastic_waste <- plastic_waste |> 
  mutate(
    coastal_population_proportion = coastal_pop / total_pop) |> 
  filter(plastic_waste_per_cap < 3)

ggplot(new_plastic_waste) +
  aes(x = coastal_population_proportion,
      y = plastic_waste_per_cap,
      color = continent) +
  geom_point() +
  geom_smooth(aes(group = 1)) +
  labs(x = "Coastal population proportion (Coastal / total population",
       y = "Plastic waste per capita",
       title = "Plastic waste vs. coastal population proportion by continent")
  
```

## Pro-Tips

### Excercise 3

Try this :D

```
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()+
  geom_boxplot(width=.3, fill="green") +
  stat_summary(fun.y=median, geom="point") 
```

### Exercise 5 

Helpful reference:http://www.sthda.com/english/wiki/ggplot2-themes-and-background-colors-the-3-elements



