---
title: 'Lab 4: Personality and Green Reputation'
author: "Reilly Orman"
date: "2/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("ggplot2")
library("dplyr")
dictionary <- readr::read_csv(here::here("data", "green_dictionary.csv"), na = c(""))
green_data <- readr::read_csv(here::here("data", "green_data.csv"))
```

1. Inspect the item responses. Is anything unusual?

```{r 1.0}
#take a look at data to inspect any unusual values
glimpse(dictionary)

glimpse(green_data)
```

```{r}
#use unique function to pull out unusual values
green_data |> 
  select(-id) |> 
  unlist() |> 
  unique()
```

```{r}
#recode data to make -99 an NA value
green_data |> 
  mutate(
    across(
      -id,
      \(x) recode(x, "1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5, "-99" = NA_real_)
    )
  )
```


```{r}
green_data <- green_data |> 
  mutate(
    across(
      -id,
      \(x) na_if(x, -99)
    )
  )
```


```{r recode}
#assigning reverse scaled columns to object 
dict <- dictionary |> 
  filter(Keying < 0) |> 
  pull(Item) 
```

```{r}
#rescale reversed scaled variables
new_green_data <- green_data |> 
  mutate(
    across(
      all_of(dict), 
      \(x) recode(x, "1" = 5, "2" = 4, "3" = 3, "4" = 2, "5" = 1)
      ))
```

```{r total scores}
#computing total scores for the four scales
new_green_data <- new_green_data |>
  rowwise() |> 
  mutate(
    green_total = mean(c_across(green1:green5), na.rm = TRUE),
    comp_total = mean(c_across(comp1:comp10), na.rm = TRUE),
    intel_total = mean(c_across(intel1:intel10), na.rm = TRUE),
    open_total = mean(c_across(open1:open10), na.rm = TRUE),
    .before = everything()
    ) |> 
  ungroup() 
  
print(new_green_data)

```

3. Rescale the variables so they go from 0-100 instead of the original range. Name the rescaled variables *_pomp. 

```{r rescale}
dat_green_data <- new_green_data |> 
  mutate(
    across(
      green_total:open_total,
      \(x) 100 * (x-1) / (5-1),
      .names = "{.col}_pomp"
    ),
    .before = id
  )
```

4. Plots that illustrate the distributions of the 4 POMP-scored variables. 

```{r}
#histogram for green total pomp
ggplot(dat_green_data) +
  aes(x = green_total_pomp) +
  geom_histogram()
```

```{r}
#histogram for comp total pomp
ggplot(dat_green_data) +
  aes(x = comp_total_pomp) +
  geom_histogram()
```

```{r}
#histogram for intel total pomp
ggplot(dat_green_data) +
  aes(x = intel_total_pomp) +
  geom_histogram()
```

```{r}
#histogram for open total pomp
ggplot(dat_green_data) +
  aes(x = open_total_pomp) +
  geom_histogram()
```

5. Make scatterplots showing the relationships between green reputation and each personality trait. Include trend lines for students and non-students. What do these plots show?

```{r}
#scatterplot showing relationship between compassion and green reputation
dat_green_data |>
  filter(!is.na(student)) |> 
  ggplot() +
  aes(x = comp_total_pomp,
      y = green_total_pomp,
      color = factor(student)) +
  geom_point() +
  geom_smooth(method = "lm")
```


```{r}
#scatterplot showing relationship between intellectual curiosity and green reputation
dat_green_data |>
  filter(!is.na(student)) |> 
  ggplot() +
  aes(x = intel_total_pomp,
      y = green_total_pomp,
      color = factor(student)) +
  geom_point() +
  geom_smooth(method = "lm")
```


```{r}
#scatterplot between green reputation and openness to experience personality trait
dat_green_data |>
  filter(!is.na(student)) |> 
  ggplot() +
  aes(x = open_total_pomp,
      y = green_total_pomp,
      color = factor(student)) +
  geom_point() +
  geom_smooth(method = "lm")
```

There appears to be a somewhat positive correlation between these personality traits and green reputation. Openness to experience seems to have the strongest relationship based on the trend line. 

6. Compare green reputation for students and non-students using a rainfall plot

```{r}
#recode student variable from numeric values to character values
dat_green_data <- dat_green_data |> 
  mutate(
    dat_green_data, Student = factor(student, levels = c("1", "2"), labels = c("no", "yes")))

#rainbow plot
dat_green_data |> 
  filter(!is.na(Student)) |> 
  ggplot() +
  aes(x = green_total_pomp,
      y = Student,
      color = Student, fill = Student) +
  geom_jitter(height = .15) +
  geom_boxplot(color = "black",
               alpha = .5,
               width = .1,
               size = .5) +
  ggdist::stat_slab(height = .3,
                    color = "black",
                    size = .2,
                    alpha = .5,
                    position = position_nudge(y = .2))
```

7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.

```{r}
#summary table for total scores 
dat_green_data |> 
  group_by(Student) |> 
  summarize(across(green_total_pomp:open_total_pomp,
                   mu = mean(),
                   sigma = sd(),
                   median = median(),
                   min = min(),
                   max = max()))
```

